---
title: "scratch"
output: html_document
date: "2025-10-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
```

```{r cars}
# load dataframe
all_credit <- read_csv(here::here("Data", "application_train.csv"))
```

```{r}
nrow(all_credit)
```


```{r}
# aggregate documents flagged, one-hot-encode, and select columns, 
credit <- all_credit %>% 
  mutate(
    DOCUMENTS_FLAGGED = rowSums(select(., starts_with("FLAG_DOCUMENT")))
  ) %>% 
  mutate(
    CONTRACT_TYPE_CASH = ifelse(NAME_CONTRACT_TYPE=="Cash loans", 1, 0), 
    CONTRACT_TYPE_REVOLVING = ifelse(NAME_CONTRACT_TYPE=="Revolving loans", 1, 0),
    EDUCATION_TYPE_SECONDARY = ifelse(NAME_EDUCATION_TYPE=="Secondary / secondary special", 1, 0),
    EDUCATION_TYPE_HIGHER = ifelse(NAME_EDUCATION_TYPE=="Higher education", 1, 0), 
    EDUCATION_TYPE_OTHER = ifelse(!NAME_EDUCATION_TYPE %in% c("Secondary / secondary special", "Higher education"), 1, 0),
    FAMILY_STATUS_MARRIED = ifelse(NAME_FAMILY_STATUS=="Married", 1, 0), 
    FAMILY_STATUS_SINGLE = ifelse(NAME_FAMILY_STATUS=="Single / not married", 1, 0), 
    FAMILY_STATUS_OTHER = ifelse(!NAME_FAMILY_STATUS %in% c("Married", "Single / not married"), 1, 0),
    HOUSING_TYPE_OWN = ifelse(NAME_HOUSING_TYPE=="House / apartment", 1, 0), 
    HOUSING_TYPE_PARENTS = ifelse(NAME_HOUSING_TYPE=="With parents", 1, 0), 
    HOUSING_TYPE_OTHER = ifelse(!NAME_HOUSING_TYPE %in% c("House / apartment", "With parents"), 1, 0), 
    FLAG_OWN_CAR = ifelse(FLAG_OWN_CAR =="Y", 1, 0),
    FLAG_OWN_REALTY = ifelse(FLAG_OWN_REALTY=="Y", 1, 0)
        ) %>%
  select (
    SK_ID_CURR,
    CONTRACT_TYPE_CASH, 
    CONTRACT_TYPE_REVOLVING, 
    FLAG_OWN_CAR,
    FLAG_OWN_REALTY,
    CNT_CHILDREN,
    AMT_INCOME_TOTAL,
    AMT_CREDIT,
    AMT_ANNUITY,
    AMT_GOODS_PRICE,
    EDUCATION_TYPE_SECONDARY, 
    EDUCATION_TYPE_HIGHER, 
    EDUCATION_TYPE_OTHER, 
    FAMILY_STATUS_MARRIED, 
    FAMILY_STATUS_SINGLE, 
    FAMILY_STATUS_OTHER, 
    HOUSING_TYPE_OWN, 
    HOUSING_TYPE_PARENTS, 
    HOUSING_TYPE_OTHER,
    DAYS_EMPLOYED,
    DOCUMENTS_FLAGGED, 
    TARGET
  )

head(credit)
```
```{r}
nrow(credit)
```

LIST OF (new) VARIABLES:

- SK_ID_CURR
- TARGET
- NAME_CONTRACT_TYPE
  -> CONTRACT_TYPE_CASH
  -> CONTRACT_TYPE_REVOLVING
- FLAG_OWN_CAR
- FLAG_OWN_REALTY
- CNT_CHILDREN
- AMT_INCOME_TOTAL
- AMT_CREDIT
- AMT_ANNUITY
- AMT_GOODS_PRICE
- NAME_INCOME_TYPE? 
- NAME_EDUCATION_TYPE
  -> EDUCATION_TYPE_SECONDARY
  -> EDUCATION_TYPE_HIGHER
  -> EDUCATION_TYPE_OTHER
- NAME_FAMILY_STATUS
  -> FAMILY_STATUS_MARRIED
  -> FAMILY_STATUS_SINGLE
  -> FAMILY_STATUS_OTHER
- NAME_HOUSING_TYPE
  -> HOUSING_TYPE_OWN (house / apartment)
  -> HOUSING_TYPE_PARENTS
  -> HOUSING_TYPE_OTHER
- DAYS_EMPLOYED (current job - proxy for age?)
- DOCUMENTS_FLAGGED (aggregate of FLAG_DOCUMENT 2-21)

```{r}
plot_children <- credit |>
  group_by(CNT_CHILDREN) |>
  summarise(default_rate = mean(TARGET)) |>
  ggplot(aes(x = factor(CNT_CHILDREN), y = default_rate, fill = factor(CNT_CHILDREN))) +
  geom_col() +
  labs(title="Default Rate by Number of Children",
       x = "Number of Children",
       y = "Default Rate") +
  theme_bw() +
  theme(legend.position = "none")

plot_children
```

```{r}
# save plot locally, dimensions chosen arbitrarily
ggsave("images/default_children.png", plot=plot_children, width = 10, height = 6)
```

```{r}
plot_education <- credit |>
  mutate(
    Education = factor(case_when(
      EDUCATION_TYPE_HIGHER == 1 ~ "Higher",
      EDUCATION_TYPE_SECONDARY == 1 ~ "Secondary",
      TRUE ~ "Other"
    ), levels = c("Higher", "Secondary", "Other"))
  ) |>
  group_by(Education) |>
  summarise(default_rate = mean(TARGET)) |>
  ggplot(aes(x = Education, y = default_rate, fill = Education)) +
  geom_col() +
  labs(title = "Default Rate by Education Level",
       y = "Default Rate") +
  theme_bw() +
  theme(legend.position = "none")

plot_education
```

```{r}
# save plot locally, dimensions chosen arbitrarily
ggsave("images/default_education.png", plot=plot_education, width = 10, height = 6)
```

```{r}

# function to determine nan values, given dataframe 
# consumes dataframe, returns df of colnames and number nan

# add percent nan function also
count_nan <- function(df) {
  nan_cols <- data.frame()
  
  for (i in 1:ncol(df)) {
    name <- colnames(df[i])
    num_nan <- sum(is.na(df[i]))
    percent <- round(num_nan / ncol(df), 2)
    nan_cols <- rbind(nan_cols, data.frame("column name" = name, "nan values" = num_nan, "percent nan" = percent))
  }
  
  return(nan_cols)
}
```

```{r}
# critera: 
# keep all columns with 0 N/A
# all columns with <= 1% NA: drop those rows
# other columns: remove from the model (too many NA)

# QUESTION: can we make assumptions about WHY some values might be Nan? 

```

```{r}
count_nan(all_credit)

count_nan(credit)
```
